---
title: "Accessing USDA CRP Enrollment Data"
author: "Andrew Leonard"
date: "7/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

### This R Markdown document will provide an example of how to access data for the [USDA's Conservation Reserve Program (CRP)](https://www.fsa.usda.gov/programs-and-services/conservation-programs/conservation-reserve-program/).  In this example, I will read an Excel document located on the USDA's website containing data for [CRP Enrollment and Rental Payments by State, 1986-2019](https://www.fsa.usda.gov/Assets/USDA-FSA-Public/usdafiles/Conservation/Excel/HistoryState86-19.xlsx).  Then I will clean this data and save it locally to a CSV.

Note, if a variable name begins with a `.` this is simply to avoid conflicting names with [R keywords](https://rdrr.io/r/base/Reserved.html) or other variables.

### Read

Let's start by loading the necessary packages
```{r}
library("openxlsx")
library("data.table")
library("ggplot2")
suppressMessages(library("here"))
```

Read data
```{r}
usda_url <- "https://www.fsa.usda.gov/Assets/USDA-FSA-Public/usdafiles/Conservation/Excel/HistoryState86-19.xlsx"
DT <- read.xlsx(usda_url, sheet = "ACRES", startRow = 4)
setDT(DT)
```

View the top 6 rows of the table.
```{r}
head(DT)
```

### Clean 

Un-pivot the data in order to make it easier to work with.
```{r}
DT <- data.table::melt(DT, id.vars = "STATE", variable.name = "YEAR", value.name = "ACRES")
```

View the top 3 rows of the table.
```{r}
head(DT, 3)
```

View the bottom 3 rows of the table.
```{r}
tail(DT, 3)
```

Looks like we grabbed a blank row (row where STATE = 1/ Fiscal years end September 30th. Not including Virginia cities.) as well as the US total (row where STATE = U.S.).  Let's filter the entire table by selecting only rows containing the name of an actual state.
```{r}
state_names <- toupper(state.name)
DT <- DT[STATE %in% state_names]
unique(DT$STATE)
```

Format column names so they're all lowercase to make it easier to work with.
```{r}
old_names <- names(DT)
new_names <- tolower(old_names)
setnames(DT, old_names, new_names)
```

Let's aggregate the data in order to plot the annual US total.
```{r}
dt <- DT[, .(total_acres = sum(acres), avg_acres = mean(acres)), by = .(year)]
ggplot(dt, aes(x = as.character(year), y = total_acres)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45)) + 
  labs(y = "Total Acres", x = "Year")
```

Save the results.
```{r}
csv_fname <- here("Data", "usda_crp_enrollment")
fwrite(DT, csv_fname)
```

And that's it!  I hope this was helpful.  Please feel free to reach out with any questions or concerns.